### 2021-10-21 manuscript_1.0.Rmd
New set up to fit also de scaling constant. A bit of progress in the manuscript. Got things ready to run fitting procedure on simulated data. Started test to use R package to run MCMCs with any typpe of model.

---
title: "Cognitive and ecological aspects determine cleaner performance"
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia - andresqp@gmail.com]
- Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  word_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: Cleaner_learning.bib
style: C:\Users\andre\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: My abstract.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
```

## Introduction

Cooperation among individuals of the same or different species is ubiquitous in nature. 
Among individuals of the same species it is the basis for the evolution of  higher 
forms of biological organization [@maynardsmith_Major_1998]. 
Among individuals of different species it underlays the biological mechanism for 
fundamental ecological interactions such as pollination, nitrogen fixation or 
parasite removal [@begon_Ecology_2006]. However, when cooperation is costly, 
it is not straigtforward to understand why individuals would be willing to pay that cost [@lehmann_Evolution_2006]. Furthermore, cooperative interactions might open the way for 
one partner to exploit the other, what prevents this type of exploitation? [@herre_Evolution_1999]
One reason for cooperating is if the benefits obtained from a partner´s 
cooperative action are dependent on one's own cooperative act. 
This dependency can arise, for instance, when cooperative partners
choose who they interact with depending on the level of cooperation they offer. 
In line with this, cooperative interactions can be viewed as a market-like situation 
where partner choice mediates the level of exchange between traders [@hammerstein_Biological_2016; @leimar_Cooperation_2010; @noe_Biological_1994a]. Empirical evidence shows
that several iconic examples of cooperation between individuals of different species 
are indeed mediated by partner choice [@hammerstein_Biological_2016]. Biological market theory highlights 
that changes in the supply and demand of goods and services determines the value of 
those commodities, and therefore, can influence partner's actions in the cooperative 
exchange. For instance, in the marine cleaning mutualism, where cleaner fish 
_Labroides dimidiatus_ (cleaners) removes ectoparasites from other coral reef fishes (clients) [@cote_Evolution_2000],
some client species are able to choose among different cleaners, and because of this
obtain better service [@grutter_Cleaner_2003]. Better services means for clients to have priority when 
more than one client is available [@bshary_Choosy_2002], and less cheating (when cleaners eat live tissue
instead of parasites) [@grutter_Cleaner_2003]. However, recent evidence shows that when clients experience
ecological conditions with low cleaner abundance, they are less likely to use partner
choice to leverage better service [@triki_Biological_2019]. On the other side, 
cleaners in those markets with favourable conditions, stop giving priority to choosy clients [@triki_Biological_2019; @triki_Decrease_2018]. Thus, it seems that cleaners, as well as clients,
adjust their strategy to the prevalent market conditions they find in their environment. 
This adjustement cannot be due to local genetic adaptation, given that cleaners are open water spawners 
and the environmental changes in different sites have occured in the last decade [@triki_Biological_2019; @triki_Decrease_2018]. Alternatively, a recent cognitive model proposes that differences in 
cleaners preferences can arise trough the interaction of cleaner's cognitive processes 
with the environmental conditions [@quinones_Reinforcement_2019]. Under this view,
cleaners ontogenatically develop their preferences for different clients through learning,
and some environmental conditions facilitate learning. Thus, the adaptive adjustment 
seen in cleaners might be due to their learning mechanisms. Despite the general fit of the 
model explanation to the available data, no systematic parameter fit has been performed to 
evaluate which combination of model parameters better fit the data.

## Methods

In order to fit the model parameters to the data we used Markov Chain Monte Carlo (MCMC) with 
a Metropolis sampler to estimate the posterior distribution of the model parameters. We used the data from @triki_Biological_2019 to fit the parameters of the  model analyzed in @quinones_Reinforcement_2019. The data consist a of the number of times that set of cleaners chose to
feed of a client plate in the context of the market task. This set of cleaners were collected under different locations and in each one of those locations environmental variables were measured. These environmental variables recorded from each location are used as input for the computational model. Given these inputs, the computational model produces a time series of choices in the market task. In the model the preference for the two option changes as the algorithm learns based on the reward that the two options provide. Thus, we run the model for 10000 steps and use the las 3000 steps as the preference developed in the model for visitors given the environmental variables. We use this preference as the predicted probability of choosing a visitor in the data, and use it to calculate the likelihood that the number of visitor choices obtained from each cleaner is generated by the model. We sum over all the cleaners the loglikehoods and use the total for the acceptance of new parameters in the MCMC.

The model consist of a set individual-simulations where individuals face a series of choices between two options. Individuals in the model represent cleaners, and the options they face simulate the natural conditions of a cleaning market. Thus, individuals in the model choose to prioritize  different clients. These clients are classified as resident and visitors. Residents are those clients that stay in the cleaning queue when they are not given priority; while visitors leave the queue (with a certain probability) after they are not given priority. Individuals obtain reward from cleaning a client regardless of the type.  Everytime individuals face and make a choice they update the probability of making that same choice. The update is based on the difference between the expectation of reward and the actual reward the individual obtains (prediction error); and it is carried in the direction that would lead to more reward given the new information. Further details of the model implementation can be found in the paper [@quinones_Reinforcement_2019]. The model analysis points to the relevance of two parameters. First, $\gamma$ measures how much individuals include future rewards in their desicion updates. If $\gamma=0$, individuals only use the immediate reward obtained from a cleaning interaction. As $\gamma$ increases individuals include more the reward obtained from the subsequent choices. That amounts to calculating and using for deision making future expected rewards. Second, $\eta$ measures how much individuals include in their reward the fleeing behaviour of visitor clients as a negative component. Both of these components allow individuals to use in ther estimates the future effects of their choices. The model analysis shows that without either of them, learning individuals cannot develop a preference for the visitor client. The analysis contrast the effect of these two parameters agaisnt variation in the relative abundance of the two client types. The analysis shows that under very high client abundance the presence of  future rewards ($\gamma>0$) no longer favours the preference for the ephemeral clients, this effect does not occur with negative reward mediated by client behaviour ($\eta$). Thus, the analysis suggests that future reward as an explanation for cleaner ability to prefer visitors is supported by @triki_Biological_2019 data. However, a combination of future and negative rewards, could still be a better explanation for the data. In the current analysis, we will use the data from @triki_Biological_2019 to fit some of the model parameters, particularly relevant we will fit $\gamma$ and $\eta$ to find out whether each or a combination of these effects is a better explanation for the pattern seen in the data.

### Data explanation

*Insert here brief description of the data and data collection*

In order to capture with the model the relationship between the characteristics of the market (environment) and the cleaner's performance we need to translate the absolute abundances of the data to relative abundances. This is because in the model, relative abundances of clients define not only the probability of residents and visitors, but also how often the cleaning station is empty (e.g. there are no clients to be cleaned). The more cleaners, the more options available for the clients and hence higher probability of empty cleaning stations. How often the cleaning stations are empty in turns influences the value that cleaners can expect from their choices. However, the frequency of empty spots in the cleaning station does not only depend on cleaner relative abundances but also on the frequency with which clients visit the station. The client visitation frequency will likely vary among different clients depnding on parasite abundances and loads. There are variables for which we do not have field estimates. Thus, we use
computed a measure of relative cleaner abundance for each site relative to absolute abundances and multiplied by scalling constant that changes the range of the variable. We fit the value os the scalling constant as part of the statistical inference. As for the client abundances for residents and visitors we computed a relative measure with respect to the total client abundance, and weigthted that by the rescaled relative absence of cleaners (1-relative cleaner abundance). Thus, all three measures of relative abundance sum up to 1, and can be used in the model as a proxy for the probability of having different options in the cleaning station.

### Results

```{r, fig1}
knitr::include_graphics(here("Simulations","ABCclean_gam_nRew_sca_",                            "round_gamma_0.3regRew_0.11scaC_133.5alphA_0.05alphC_0.05_contour_ggplot.png"))
```

## Cited literature