### 2021-11-16 manuscript_1.0.Rmd
started  with the dicussion in the manuscript. Used functions to make the data loading more straightforwd

---
title: "Cognitive and ecological aspects determine cleaner performance"
author:
- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
  - andresqp@gmail.com]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: Cleaner_learning.bib
style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: Insert a nice abstract here.
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
```

## Introduction

Cooperation among individuals of the same or different species is
ubiquitous in nature. Among individuals of different species it
underlays the biological mechanism for fundamental ecological
interactions such as pollination, nitrogen fixation or parasite removal
[@begon_Ecology_2006]. When partners in these mutualistic interactions 
are able to choose their interaction partner, their investment and the return 
they obtain can be understood as a 
biological market [@hammerstein_Biological_2016; @leimar_Cooperation_2010;
@noe_Biological_1994a,@noe_Biological_1995a]. Thus, we expect supply 
and demand of the good 
and services being exchanged to determine the action of individuals, at least in
an evolutionary time-scale.
However, supply and demand conditions of a biological market, which typically
depend on the abundance of the species involved, certainly vary in time and space. 
How do partners adjust their strategy in the face of this variation? 
In animals an obvious general candidate mechanism for the strategic adjustment is the 
cognitive machinery. However, it is not clear which specific cognitive mechanisms
allow individual to adjust their behaviour to the varying market conditions.

One example of strategic adjustment in a biological market is the marine cleaning 
mutualism involving the cleaner
fish *Labroides dimidiatus* and 'client' fish. Clients offer themselves as
food patches for cleaners to remove ectoparasites, which gives the
cleaners calories and the clients improved health
[@waldie_LongTerm_2011; @ros_Does_2011,@triki_Effects_2016]. As
cleaners have about 2000 interactions per day
[@grutter_Relationship_1995], they regularly face the situation that
two clients seek cleaning simultaneously. In such conditions, it is
important to distinguish between 'resident' and 'visitor' client species
[@bshary_Cleaner_2001a]. The residents have small home ranges and hence
access to only the local cleaner, while the visitors have large home
ranges and hence access to two or more cleaners. Thus, visitors have the
option to switch to another cleaner if being made to wait, while
residents must wait for inspection. Indeed, visitors have been observed
to use their partner choice options in that way [@bshary_Choosy_2002],
which may explain why cleaners give visitors service priority in a field
study in the Red Sea [@bshary_Cleaner_2001a]. Furthermore, early
laboratory experiments showed that leaving causes service priority
[@bshary_Asymmetric_2002]. The experiments consisted of a simultaneous
two-choice task, where cleaners had to decide which of two Plexiglas
plates to approach first and eat mashed prawn of it. One plate was made
to behave like a resident, i.e. it would remain until the cleaner had
eaten off it. The other plate was made to behave like a visitor, i.e. it
would be removed if not inspected first. Cleaners learned to prefer the
visitor plate and hence obtained the double amount of food
[@bshary_Asymmetric_2002]. Adult cleaner fish outperform various
primates as well as rats and pigeons in this original version of the
biological market task [@salwiczek_Adult_2012; @zentall_Early_2017];
only African grey parrots solve this task as well [@pepperberg_Can_2014].

Recent research has refined our view on partner choice and priority of
access in the *L. dimidiatus* cleaning mutualism, and as a consequence
performance in the biological market task. First, details of the task
presentation may matter; i.e. what relevant information is encoded in
the design of the two plates (colour versus size; @wismer_Cuebased_2019), or
whether the information is in the colour of the food
[@pretot_Comparative_2021; @pretot_Comparing_2016], and whether the
food is visible or hidden/released only when the subjects made choices
on a touch screen [@pretot_Factors_2016]. Second, it appears that cleaner and
client densities in nature affect how cleaners perform in the original
market task: when densities are low, cleaners are more likely to fail to
complete the task [@triki_Biological_2019; @triki_Decrease_2018;
@wismer_Variation_2014]. The apparent explanation for this failure is that visitor
clients are less likely to exert their choice options (and hence wait
for inspection if made to wait) when cleaner fish densities are low
[@triki_Brain_2020]. This change in behaviour may be due to the fact that
when cleaner densities are low, the cost of switching in terms of
distance to swim and time invested increases. If visitors are not
choosy, cleaners should not give them priority anymore. Interestingly,
while cleaners from high-density areas have larger forebrains harbouring
more cells than cleaners from low-density areas [@triki_Brain_2020], on a
local scale those individuals that perform in the market task in a way
that corresponds to the locally optimal strategy (solving/failing the
task when living in high/low density) have relatively larger forebrains
[@triki_Brain_2020]. Taken together, recent empirical results clearly show
that cognitive processes are involved in the pattern of intra-
and inter-specific variation in  performance in the market task. One approach to
identify candidate mechanisms underlying performance in a cognitive task
is to develop proximate learning models and to evaluated in light of the behavioural data 
(e.g. @dubois_Model_2021). By varying the cognitive tool
kit, one can assess which minimal kit is necessary to solve the task at
hand. This approach has been used to analyse the market task
[@prat_Role_2021; @quinones_Reinforcement_2019]. Learning models showed that basic reinforcement learning does not suffice to solve the task, this is 
particularly so when model assume the more complex natural situation 
in which cleaners
not only face resident-visitor pairs but also visitor-visitor and
resident-resident pairs, as well only a resident or only a visitor. To
be able to give visitors priority over residents, cleaners first need to
be able to assess a client's value separately for the three possible
forms of encounter (alone, paired with a fish with the same strategic
option, paired with a fish with the alternative strategic option)
[@quinones_Reinforcement_2019]. In addition, cleaners need to account for the future
consequences of current decisions. In the model, this could be achieved
in two non-mutually exclusive ways [@quinones_Reinforcement_2019]:
through low temporal discounting of future effects
and/or through perceiving a visitor client leaving as psychological
punishment (i.e. as a negative reinforcer). Given that a negative
reinforcer is part of the ubiquitous learning mechanism termed operant
conditioning (Thorndike xxxx), this explanation can be considered the
simpler one. Furthermore, accounting for the future effects of an action is 
considered an advanced cognitive ability [@suddendorf_Evolution_2007]. 
Here, we fitted our model to the empirical data on fish
densities and cleaner fish performance in the market task with the aim
to determine the likely mechanism used by cleaners to incorporate future
consequences of current decisions.

## Methods

### Overview

We used the data from @triki_Biological_2019 to fit the parameters of the model analysed in
@quinones_Reinforcement_2019. For the bayesian fitting procedure we used Markov Chain Monte Carlo (MCMC) with a Metropolis sampler to estimate the posterior
distribution of the model parameters.  The data consist a of the number of times
that a set of cleaners chose to feed of a visitor client plate in the
context of the market task. This set of cleaners were collected in
different locations and in each one of those locations environmental
variables were measured. These environmental variables recorded from
each location are used as input for the computational model. Given these
inputs, the computational model produces a time series of choices in the
market task. In the model the preference for the two options changes as
the algorithm learns based on the reward that the two options provide.
Thus, we run the model for 10000 steps and use the last 3000 steps as
the preference developed in the model for visitors given the
environmental variables. We use this preference as the predicted
probability in a binomial distribution to calculate
the likelihood that the number of visitor choices obtained from each
cleaner is generated by the model. We sum over all the cleaners the
log-likehoods and use the total for the acceptance of new parameters in
the MCMC.

### The model

The model consist of a set individual-based simulations where
individuals face a series of choices between two options. Individuals in
the model represent cleaners, and the options they face simulate the
natural conditions of a cleaning market. Thus, individuals in the model
choose to prioritize different clients. These clients are classified as
resident and visitors. Residents are those clients that stay in the
cleaning queue when they are not given priority; while visitors leave
the queue (with a certain probability) after they are not given
priority. Individuals obtain reward from cleaning a client regardless of
the type. Every time individuals face and make a choice they update the
probability of making that same choice. The update is based on the
difference between the expectation of reward and the actual reward the
individual obtains (prediction error); and it is carried in the
direction that would lead to more reward being obtained, given the new
information. Further details of the model implementation can be found in
the paper [@quinones_Reinforcement_2019]. The model analysis points to
the relevance of two parameters. First, $\gamma$ measures how much
individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards in action. Second,
$\eta$ measures how much individuals include in their reward the fleeing
behaviour of visitor clients as a negative component. Both of these
parameters allow individuals to use in their estimates the future
effects of their choices. The model analysis shows that without either
of them, learning individuals cannot develop a preference for the
visitor client. The analysis contrast the effect of these two parameters
against variation in the relative abundance of the two client types. The
analysis shows that under very high client (relative) abundance the presence of
future rewards ($\gamma>0$) no longer favours the preference for the
ephemeral clients, this effect does not occur with negative reward
mediated by client behaviour ($\eta$). Thus, the analysis suggests that
future reward as an explanation for cleaner ability to prefer visitors
is supported by @triki_Biological_2019 data. However, a combination of
future and negative rewards, could still be a better explanation for the
data. In the current analysis, we will use the data from
@triki_Biological_2019 to fit some of the model parameters, particularly
relevant we will fit $\gamma$ and $\eta$ to find out whether each or a
combination of these effects is a better explanation for the pattern
seen in the data.

### Data explanation

*Insert here brief description of the data and data collection*

In order to capture with the model the relationship between the
characteristics of the market (environment) and the cleaner's
performance we need to translate the absolute abundances of the data to
relative abundances. This is because in the model, relative abundances
of clients define not only the probability of residents and visitors,
but also how often the cleaning station is empty (e.g. there are no
clients to be cleaned). The more cleaners, the more options available
for the clients and hence higher probability of empty cleaning stations.
How often the cleaning stations are empty in turns influences the value
that cleaners can expect from their choices. The more empty the empty
cleaning stations are, the more it is to choose a client that will not
stay. However, the frequency of empty spots in the cleaning station does
not only depend on cleaner relative abundances but also on the frequency
with which clients visit the station. The client visitation frequency
will likely vary among different clients depending on parasite
abundances and loads. These are variables for which we do not have field
estimates. Thus, we computed a measure of relative cleaner abundance for
each site relative to absolute abundances and multiplied by a scaling
constant that changes the range of the variable. We fit the value of the
scaling constant as part of the statistical inference. As for the client
abundances for residents and visitors we computed a relative measure
with respect to the total client abundance, and weighted that by the
rescaled relative absence of cleaners (1-relative cleaner abundance).
Thus, all three measures of relative abundance sum up to 1, and can be
used in the model as a proxy for the probability of having different
options in the cleaning station.

## Results

```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scaling constant. We show the kernel density estimates, below the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres. On the left (panels a, c and c) posterior distributions for a model with   all three parameters; while on the right (b and d) a model without negative reward. Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='85%'}
knitr::include_graphics(here("post_both_gamma.png"))

```

Figure \ref{fig:post} panels a,c and e show the posterior distributions
of the parameter values of the full model, which includes both future
and negative reward. Notably, the bulk of the posterior distribution for
the parameter tuning negative reward ($\eta$ panerl e) is around zero. Thus, the shape of the posterior and its credible intervals suggest the absence of negative reward to be well supported by the data. As for $\gamma$, the confidence intervals also includes zero, but the mode of the posterior is around 0.5. Thus, we run the analysis setting $\eta$ to zero. Figure
\ref{fig:post} panels b and d, show the posterior distributions of the
two parameters estimated. Not surprisingly, under a model without
negative reward, the distribution of $\gamma$ shifts to higher values. Under the new model, the estimates of $\gamma$ do not include zero, the distribution is centered arounf $0.5$.
To asses the fit of the two models we calculated the  $pseudo-R^2$ [@mcfadden_Conditional_1974], which is a standard measure of fit for logistic regression. 
In that context, $pseudo-R^2$ uses the log-likelihood of the data given the model, relative to the log-likelihood of the data given a model without covariates, as a measure of fit. 
Given that our model is not a logistic regression, we measure the $pseudo-R^2$ 
in relation to the log-likelihood of a model with parameters $\gamma$ and $\eta$ set to zero. This in practice amounts to a model that has neutral preference between 
the two options. In panel f of figure \ref{fig:post} we show the distribution of $pseudo-R^2$ calculated using samples from the posterior distributions shown before, as well as from a 
model with negative reward but without future reward. 
Note, $pseudo-R^2$ can have negative values, that is when the log-likelihood of 
the model is lower than that of a model that triggers neutral preferences. 
Even though, the peak of the three $pseudo-R^2$ distributions is not very different, 
the model with only future reward produces a distribution of $pseudo-R^2$ where more values 
are positive (to the right of black line in Fig. \ref{fig:post} f). This 
shows that, accounting for variation in the parameter estimates, the model with only future
reward gives a better fit to the data, despite having one parameter less.

```{r pred, fig.cap="Observed and predicted probability of choosing a visitor. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panels: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities. Pseudo $R^2$ shown in the lower right corner corresponds to the fit of a model parameterized with the mode of the posterior distributions. Upper panels show prediction from a model including future and negative reward, lower panels froma  model with only future reward.",out.width='90%'}

knitr::include_graphics(here("contour_both_gamma.png"))

```

Figure \ref{fig:pred} shows predictions from the two models. 
On the left, the colour coded contour shows the
prediction of the model for combinations of cleaner abundance (x axis)
and visitor leaving probability (y axis). Predictions were generated by
running the model using the mode of the posterior distributions as the
parameter values, and varying systematically cleaner abundance and
visitor leaving probability. Colour of the points on top of the contour
show the average frequency with which cleaner fish chose the visitor
plate in their last 20 experimental trials. Each point corresponds to a
location with a combination of cleaner abundance and visitor leaving
probability. Panels on the right show how close the observed values for 
the probability of choosing the visitor (y
axis) are from the predictions of the models (x axis). The line
represents the perfect fit between observed and predicted values.
Predicted values were taken using the mode just as in the left panels.
On the lower right corner of the panel we show the $pseudo-R^2$ for predictions
using the mode of the posterior distributions. 
Despite having one more parameter varying, the full model
gives a lower fit to the data than the model without negative reward. 
The difference in $pseudo-R^2$ for the point estimates is small, but using the whole posterior distribution
shows the improvement in fit achieved by eliminating negative reward 
from the model (Fig. \ref{fig:post}f). In the supplementary
material, we also show that a model with only negative reward, excluding
future rewards ($\gamma=0$, Fig. \ref{fig:nogamma}), provides a lower
fit than either of the ones presented in the main text.

The main reason for negative and future reward to give different
predictions is the way that cleaner's relative abundance influences the
preference for the visitor clients. Visitor leaving probability has a
similar positive effect on the probability of choosing the visitor
clients on all three models (Fig. \ref{fig:pred}, \ref{fig:nogamma}). In
contrast, cleaner's relative abundance has a different effect in the
model with only future reward, compared to the other two. In the model
with only future reward only intermediate cleaner abundance triggers a
preference for the visitor clients (Fig. \ref{fig:pred} c). In models
with negative reward, both intermediate and low cleaner abundance
trigger a preference for the visitor. (Fig. \ref{fig:pred} a,
\ref{fig:nogamma}).

## Discussion

We have shown that using future rewards in the learning process is the cognitive 
strategy that best explains the intra-specific variation in cleaner performance 
in the market task. We compared future reward with the use of client behaviour as 
a primary source of negative reward, and a combination of both. All three models 
capture well the positive relation between visitor leaving behaviour and cleaner
performance in the market task [@triki_Biological_2019]. For both mechanisms it is 
indeed necessary to have fleeing behaviour on the side of the visitors for cleaners
to develop a preference for them. The key difference between future and negative 
reward emerges in the predicted relation between cleaner relative abundance and 
cleaner performance. When cleaner relative abundance is low a model with only 
future reward predicts low performance in the market task regardless of the visitor
leaving probability. In contrast, A model with negative reward predicts the highest 
performance with low relative cleaner abundance, particularly together with high 
probability of visitor leaving (Fig. \ref{fig:pred}). Low cleaner abundances 
mean the market has an excess of demand for cleaning services. In the model,
this translates to a cleaning station that is frequently full. Thus, when visitors 
leave, it's likely that the cleaner will have access to another client, and there won't be much difference in future reward between choosing a visitor and a resident. Thus,
estimation of future reward will not develop a preference for the visitor in this 
conditions, because it matches the state of the market. On the other hand, the effect of negative reward is unaffected by this market situation. In a full station, cleaners will get more often the heterotypic (resident-visitor) option and will develop a preference for the visitor faster. Estimation of future reward is then the cognitive mechanism 
that allows cleaner to adaptively respond to the ecological conditions of the biological market.

<!-- ### Correlation of visitor leaving and cleaner density  -->

<!-- ### Cognition as an explanation -->

<!-- bash on social competence -->

<!-- Cognitive processes gain support as an explanation for inter and intra-specific  -->
<!-- variation in the market task. Even though the model considered here does not  -->
<!-- explicitly model perception, associative processes must include the perception  -->
<!-- of different stimuli which allow cleaners to discriminate -->
<!-- residents and visitors. Thus, the fact that recent experimental work shows -->
<!-- an effect of the perceptual conditions in the experiments on the performance of  -->
<!-- the subjects, supports the idea that learning is involved in the decision  -->
<!-- making process of the market task. Furthermore, and notably, certain features  -->
<!-- of brain anatomy correlate with performance in the market task and with the  -->
<!-- abundance of cleaners in the location where the fish were collected [@triki_Brain_2020].  -->
<!-- All this, supports the idea that learning processes are behind the behavioural  -->
<!-- adjustment that cleaners show to environmental conditions.  -->

## Cited literature

::: {#refs}
:::

\newpage

# Supplementary material

```{=tex}
\beginsupplement
```
```{r nogamma, fig.cap="Observed and predicted probability of choosing a visitor for a model without future rewards. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panel: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities.",out.width='90%'}
fileNogamma<-grep("contour_gamma_",list.files(here("Simulations","ABCclean_nRew_sca_"),
                                              full.names = TRUE),value = TRUE)
knitr::include_graphics(fileNogamma)

```