### 2022-05-26 prog_report/report_attention.Rmd
Preliminary results on the implementation of Mackintosh and pearce and hall to the market experiment. PRogress report.

---
title: "The role of attention in learning to solve the ephemeral reward task"
author: "Andres Quiñones"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:  
  bookdown::pdf_document2:
     number_sections : false 
     toc: false 
bibliography: "../manuscript_1.0/Cleanerlearning.bib"
keep_tex: yes
header-includes:
  - \usepackage{amsmath}
---

```{r setup,include=FALSE,}
knitr::opts_chunk$set(echo = FALSE,message = FALSE)
```

## The model

Just like like in previous versions of the model, we assume that clients that 
reach the cleaning station of a cleaner fish can be classified into resident 
or visitors. Up to two of these clients can demand cleaning service simultaneously.
In such case the cleaner fish must choose which client to clean first. In order to
make that choice, the cleaner fish estimates a value for each client combination
that it faces. Hence, each client combination is a **state** in Reinforcement 
Learning (RL) jargon. In order to update the estimates and use the estimations 
based on the client types, cleaners must discriminate between 
the two types and their combinations in the cleaning stations. We, thus, assume
that each client type is associated with a specific stimulus. The cleaner fish
stores an estimate of value for each of these stimuli in a look-up table 
\ref{tab:lookup}. These estimates of individual client values are used independently
or combined depending on the state of the cleaning station. For example, 
when the cleaning station only has a resident the estimated value is 
$S_r$ of the state; while when it has a resident and a visitor the 
estimated value is $S_r+S_v$. This combination of stimuli to estimate 
value is in line with classical accounts of associative 
learning [@rescorla_Theory_1972].

The estimation and decision making are structured as in the **actor-critic** 
implementation in RL. Both the value estimation update (the critic) and the 
decision making update (the actor), change proportionally to the 
prediction error ($\delta$)

$$
\delta^t = R^t + \gamma \sum_{i}^{n^{t+1}} S_i^{t+1} - \sum_{i}^{n^t} S_i^t ,
$$
where $t$ and $t+1$ refer to the current and future cleaning 
interactions respectively; $R$ is the primary reward obtained from the 
choice made; $\gamma$ is measure of how much future reward 
influences the estimation process; and the sum terms correspond 
to the estimate of value made by summing over the client 
fish present in the corresponding cleaning
interaction. The total update of the value estimation is given by the product
of $\delta$ and $\alpha_i$, which can be interpreted as how much 
attention the cleaner fish pays to each stimuli. The fact that each stimuli 
can have a different $\alpha_i$ implies that the model implement differential 
attention. Furthermore, the actual value of $\alpha_i$ may be updated 
and depend on specific rules, which will be explained later. 
So, the update rule for the critic is given by

$$ \Delta S_i^t = \alpha^t_i \delta^t.  $$

The decision between resident and visitor is implemented as a probability, 
which is given by

$$ \pi_i = \frac{1}{1+e^{-(\theta_i-\theta_j)}}$$
where $\theta_i$ and $\theta_j$ are the preferences for each of the two 
available options. These preferences start on zero and are updated on every 
cleaning interaction where the cleaner fish chooses among of the two
options according to

$$ \Delta (\theta_i-\theta_j) = \beta_i^t \delta_t2(1-\pi_i)$$
where $\beta$ is the speed of the actor update and can be also interpreted as 
the attention that is paid to the respective stimuli in the decision-making
process.

```{r lookup}
lookupTab<-data.frame(client=c("Resident","Visitor","absence"),
                      Value=c("$S_r$","$S_v$",0))
knitr::kable(lookupTab,caption = "Look up table where a cleaner fish stores the 
             estimated values for each client type.",escape=FALSE)
```

## Attention updates

In the following sections we explain three different possible rules for 
the change in the selective attention variables.

### Static attention

As a benchmark we implement first a model where attention is not selective, 
all stimuli get the same attention and this level of attention does not vary
as learning proceeds. Figure \ref{fig:bench} shows the dynamics of the 
resident *vs* visitor choices that agents make along the learning process. 
The dynamics are discriminated according to how important future reward 
is in the estimation ($\gamma$). The dynamics show that regardless of the value 
of $\gamma$ used in the simulations, this set up does not allow cleaner fish 
to develop a preference for the visitor over the resident. Interestingly, 
under $\gamma=0$ the learning process leads to a small proference fir the 
resident option; while with $\gamma>0$ agents stay with a neutral preference.

```{r bench, fig.cap="Dynamics of the resident-visitor choices under a constant and even attention to all the stimuli"}
library("jsonlite")
library("here")
library("ggplot2")
source(here("loadData.R"))
source(here("aesth_par.R"))

scenario<-"equalAttAC1";simsDir<-"Simulations"
listPar<-c("AttMech","alph");listVal<-c(0,0.01)

# param<-list(totRounds=10000,ResReward=1,VisReward=1,
#             ResProb=c(0.3),
#             VisProb=c(0.3),
#             ResProbLeav=0,VisProbLeav=1,negativeRew=-0,
#             scenario=0,
#             inbr=0,outbr=0,trainingRep=10,forRat=0.0,
#             alphaT=0.05,printGen=1,seed=1, gammaRange=I(c(0,0.5,0.8)),
#             netaRange=I(c(0)),alphaThRange=I(c(0.05)),numlearn=1,
#             propfullPrint = 0.7,
#             # alphaThNch=0.01,
#             attenMech=0, equalAttAC=TRUE,
#             folderL=paste0(here(simsDir),scenario,"_/"))
# 
# check_create.dir(here(simsDir),param = rep(scenario,1),
#                  values = c(""))
# 
# parRange<-c(1)
# 
# for(parVal in parRange){
#   param$folderL<-paste0(here("Simulations",paste0(scenario,"_")),"/")
#   param$folder<-param$folderL #paste0(folderSims,"/",listfolders[i],"/") 
#   # param$attenMech<-parVal
#   outParam<-toJSON(param,auto_unbox = TRUE,pretty = TRUE)
#   fileName<-paste("parameters_",parVal,".json",sep="") 
#   if(file.exists(paste(param$folderL,fileName,sep = ''))){
#     currFile<-fromJSON(paste(param$folderL,fileName,sep = ''))
#     if(sum(unlist(currFile)!=unlist(param))>0){
#       write(outParam,paste(param$folderL,fileName,sep = "/"))
#     }
#   }
#   else{
#     write(outParam,paste(param$folderL,fileName,sep = ""))
#   }
# }

# runSims(paste0(scenario,"_"))
# 
# system("cd Simulations \n cd staticAtt_ \n ../.././CleanAC. exe parameters:1.json")

GamVar<-do.call(rbind,lapply(
                getFilelist(here("Simulations",paste0(scenario,"_")),
                            listparam = listPar,
                            values = listVal,
                            fullNam = TRUE)$PAA,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = GamVar,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()
```

### The Mackintosh update

@mackintosh_Theory_1975 proposed a model of attentional dynamics, where the 
central idea is that stimuli which are the best predictors of reward get an
increase in the attention they enjoy during the learning process. Formally,
@mackintosh_Theory_1975 defined the attentional update as

\begin{align*}
\Delta \alpha_i &> 0 \text{ if } |\lambda - V_i | <  |\lambda - V_P| \\
\Delta \alpha_i &< 0 \text{ if } |\lambda - V_i | < |\lambda- V_P|
\end{align*}

Where $\lambda$ is the "real" associative value in a given trial, which translates
to RL jargon as the reward; $V_i$ is the associative strength of the focal stimulus;
and $V_p$ is sum of the associative strengths of the stimuli other than the focal.

Following a notation more in line with the current model, we implemented the 
@mackintosh_Theory_1975 idea as follows

$$
\alpha_i^{t+1} = \hat{\alpha}(|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-\sum_{j\neq i}^{n^{t+1}}S_j^{t}|-
|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-S_i^t|)
$$

```{r Mackintosh,fig.cap="Dynamics of the resident-visitor choices under a update inspired by @mackintosh_Theory_1975"}
library(cowplot)
scenario<-"equalAttAC1"
listPar<-c("AttMech","alph");listVal<-c(1,0.01)
list1<-getFilelist(here("Simulations",paste0(scenario,"_")),
            listparam = listPar,
            values = listVal,
            fullNam = TRUE)$PAA
TimeInterv<-do.call(rbind,lapply(
                list1,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = TimeInterv,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()

```

```{r attenDynMac,fig.cap="Dynamics of attention to the different stimuli under the model inspired by  @mackintosh_Theory_1975"}
# Load raw data
rawdynamics<-do.call(rbind,lapply(list1,loadFilewithPar,parOFint="AttMech"))

# Transform associability (\alpha) data to long format
long_alpha<-melt(rawdynamics[Age%%200==0],
                    id.vars = c("Age","Training","Gamma","AttMech"),
                    variable.name = "Client",
                    measure.vars = c("alphaRes","alphaVis"))

long_alpha[,Client:=as.factor(Client)]

ggplot(data=long_alpha,aes(x=Age,y=value,colour=Client,
                              group=as.factor(Client)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom = "pointrange")+
        geom_line(aes(group=interaction(Client,Training)),
                  alpha=0.2,size=0.2,)+
                facet_grid(Gamma~Client,scales = "free_y",
                           labeller = labeller(Client=NULL))+
        scale_color_manual(values =colorValues,name="Client",
                           labels=c("Resident","Visitor"))+
  scale_y_continuous(sec.axis = sec_axis(~ . , name = expression(gamma),
                                         breaks = NULL, labels = NULL))+
      theme_classic()+
        theme(legend.position = "top",strip.background = element_blank(),
   strip.text.x = element_blank())

```

### The Pearce and Hall

@pearce_Model_1980 proposed a model of attentional dynamics, where the 
central idea is that the level of attention is proportional to the difference
between the real associative strength and that predicted by the stimulus. 
Formally, @pearce_Model_1980 defined the attentional update as

$$
\alpha^t = \rho|\lambda^{t-1}-V^{t-1}_{\sum}| + (1-\rho)\alpha^{t-1},
$$

where $V_{\sum}^{t-1}$ represents the total associative strength triggered
by the stimulus; and $\rho$ is a constant that measure how fast are the 
jumps in attention.

Following a notation more in line with the current model, we implemented the 
@pearce_Model_1980 idea as follows

$$
\alpha_i^{t+1} = \hat{\alpha}|R^t + \gamma \sum_{j}^{n^{t+1}}S_i^{t+1}-S_i^t|
$$

```{r PandH,fig.cap="Dynamics of the resident-visitor choices under a update inspired by @pearce_Model_1980"}
scenario<-"equalAttAC1"
listPar<-c("AttMech","alph");listVal<-c(2,0.01)
list1<-getFilelist(here("Simulations",paste0(scenario,"_")),
            listparam = listPar,
            values = listVal,
            fullNam = TRUE)$PAA
GamVar<-do.call(rbind,lapply(list1,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = GamVar,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()

```

```{r attenDynPandH,fig.cap="Dynamics of attention to the different stimuli under the model inspired by  @pearce_Model_1980"}

# Load raw data
rawdynamics<-do.call(rbind,lapply(list1,loadFilewithPar,parOFint="AttMech"))

# Transform associability (\alpha) data to long format
long_alpha<-melt(rawdynamics[Age%%200==0],
                    id.vars = c("Age","Training","Gamma","AttMech"),
                    variable.name = "Client",
                    measure.vars = c("alphaRes","alphaVis"))

long_alpha[,Client:=as.factor(Client)]

ggplot(data=long_alpha,aes(x=Age,y=value,colour=Client,
                              group=as.factor(Client)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom = "pointrange")+
        geom_line(aes(group=interaction(Client,Training)),
                  alpha=0.2,size=0.2,)+
                facet_grid(Gamma~Client,scales = "free_y",
                           labeller = labeller(Client=NULL))+
        scale_color_manual(values =colorValues,name="Client",
                           labels=c("Resident","Visitor"))+
  scale_y_continuous(sec.axis = sec_axis(~ . , name = expression(gamma),
                                         breaks = NULL, labels = NULL))+
      theme_classic()+
        theme(legend.position = "top",strip.background = element_blank(),
   strip.text.x = element_blank())

```

## An alternative Mackintosh implementation

During the development of the model an alternative implementation to the @mackintosh_Theory_1975. This alternative implementation captures the original conceptual idea. However, it implements it in a slightly different way:

$$
\alpha_i^{t+1} = \hat{\alpha}(|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-\sum_{j}^{n^{t+1}}S_j^{t}|-
|R^t + \gamma \sum_{j}^{n^{t+1}} S_j^{t+1}-S_i^t|)
$$

```{r MackAltRV,fig.cap="Dynamics of the resident-visitor choices under a update inspired by a model with an alternative implementation similar to @mackintosh_Theory_1975"}
scenario<-"equalAttAC1"
listPar<-c("AttMech","alph");listVal<-c(3,0.01)
list1<-getFilelist(here("Simulations",paste0(scenario,"_")),
            listparam = listPar,
            values = listVal,
            fullNam = TRUE)$PAA
GamVar<-do.call(rbind,lapply(list1,
                file2timeInter,interV=501,parOFint="AttMech"))

ggplot(data = GamVar,aes(x=Interv,y=Prob.RV.V,group=Gamma,
                         color=as.factor(Gamma)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom ="pointrange",position = position_dodge(0.9))+
        geom_hline(yintercept = 0.5,color="red")+
        geom_line(aes(group=interaction(Gamma,Training),
                  color=as.factor(Gamma)),alpha=0.2,size=0.2,
                  position = position_dodge(0.9))+
        scale_color_manual(values =colorValues, name = expression(gamma))+
        xlab("Time")+ylab("Probability of V over R")+
        theme_classic()

```

```{r MackAltAtt,fig.cap="Dynamics of attention to the different stimuli under a model inspired by  @mackintosh_Theory_1975"}

# Load raw data
rawdynamics<-do.call(rbind,lapply(list1,loadFilewithPar,parOFint="AttMech"))

# Transform associability (\alpha) data to long format
long_alpha<-melt(rawdynamics[Age%%200==0],
                    id.vars = c("Age","Training","Gamma","AttMech"),
                    variable.name = "Client",
                    measure.vars = c("alphaRes","alphaVis"))

long_alpha[,Client:=as.factor(Client)]

ggplot(data=long_alpha,aes(x=Age,y=value,colour=Client,
                              group=as.factor(Client)))+
        stat_summary(fun.data = function(x){
                x.r<-fivenum(x)
                return(list(ymin=x.r[2],ymax=x.r[4],y=x.r[3]))
        },geom = "pointrange")+
        geom_line(aes(group=interaction(Client,Training)),
                  alpha=0.2,size=0.2,)+
                facet_grid(Gamma~Client,scales = "free_y",
                           labeller = labeller(Client=NULL))+
        scale_color_manual(values =colorValues,name="Client",
                           labels=c("Resident","Visitor"))+
  scale_y_continuous(sec.axis = sec_axis(~ . , name = expression(gamma),
                                         breaks = NULL, labels = NULL))+
      theme_classic()+
        theme(legend.position = "top",strip.background = element_blank(),
   strip.text.x = element_blank())

```

# References